---
title: "20121865_BinomialDistribution"
author: "Louise Branzuela"
date: '2022-03-10'
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### **Setting my student ID as a seed for the pseudo random number generator:**

```{r}
set.seed(20121865)
```

### **Setting the size:**

```{r}
sample(20:30,1)
```

The size for this distribution is 29


### **Setting the probability of success:**

```{r}
rbeta(1,2,2)
```

The probability of success for this distribution is 0.5817 (4 d.p.)


### **Plotting the probability mass function for the distribution:**

```{r fig.align='center'}
p = 0.5817
n = 29
plot(dbinom(0:29, n, p),  
     ylab = "Probability of Success",
     xlab = "Successes",
     main = "Probability Mass Function")
```

### **Checking if the probabilities add up to one:**

```{r}
sum(dbinom(0:29, n, p))
```

### **Using rbinom() to generate a random sample from the distribution and plotting the frequency:**

```{r}
r = rbinom(1e6, n, p)
plot(table(r),    
     ylab = "Frequency",
     xlab = "Number",
     main = "Random")
```

### **Making my own dbinom() function and comparing it to R's dbinom() function:**

```{r}
mydbinom <- function(n,r,p) {
  return(factorial(r) / (factorial(n) * factorial(r - n))* p^n * (1 - p)^(r - n))
}
```

### **Testing the function to see if the probabilities add up to one**
```{r}
sum(mydbinom(0:29, n, p))
```

### **Comparing my function to R's dbinom()**
```{r}
sprintf("%.25f", dbinom(5, n, p))
sprintf("%.25f", mydbinom(5, n, p))
```


From comparing both functions, I found that every number greater than 0 will result to both functions coming up with different probabilities in the decimals.

