---
title: "20121865_HypothesisTesting"
author: "Louise Branzuela"
date: '2022-04-06'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### **Setting my student ID as a seed for the pseudo random number generator:**

```{r}
set.seed(20121865)
```

### **Generating the mean:**
```{r}
rnorm(1,10,2)
```

### **Generating the standard deviation**
```{r}
sqrt(rchisq(1,df=2))
```

```{r}
mu = 10.32528
sd = 0.8096723
```
# **Gaussian Distribution**
### **Minimum value: (3 standard deviations to the left of the mean)**
```{r}
mu - 3 * sd
```
### **Maximum value: (3 standard deviations to the right of the mean)**
```{r}
mu + 3 * sd
```

### **Setting the sequence to x**
```{r}
minimum = mu - 3 * sd
maximum = mu + 3 * sd
x <- seq(minimum, maximum, len = 100)
```

### **Probability density function for the distribution:**
```{r}
plot(x, dnorm(x, mu, sd),
     ylab = "Probability",
     xlab = "Range",
     main = "Probability Density Function")
```
#

### **Find the probability that X lies in the range:**
```{r}
pnorm(mu + 2 * sd, mu, sd) - pnorm(mu - sd, mu, sd)
```

### **Using rnorm() to generate a random sample:**
```{r}
r = rnorm(1e6, mu, sd)
```

```{r}
hist(r,
     ylab = "Frequency",
     xlab = "Number",
     main = "Random")
```

### **Finding the steepest point of the standard normal distribution:**
Theoretically, the mean would be the highest steepest point of the
normal distribution, since the distribution is perfectly
symmetrical, it would mean the middle of the distribution would
be the steepest point of the curve which corresponds to the mean.

# **Student T-Test**
```{r}
w <- c(102.5, 102.9, 99.9, 101.9, 101.3, 98.9, 96.5, 97.7, 101.8,
       100.3)
```

### **A sensible null hypothesis:**
A sensible null hypothesis would be H~0~ = the average
weight of the bananas which is 100 grams.

### **P-value and "more extreme" in this context:**
```{r}
n = length(w)
SEM = sd(w)/sqrt(n)
1 - pnorm(mean(w), 100, SEM)
```
The p-value is defined as the likelihood of 
receiving the observation or a more extreme observation 
if the null hypothesis is true. The p-value that I got was 0.2910448,
in this dataset, "more extreme"
means larger than the observed sample mean.

### **One-sided or two-sided test:**
A one-sided test(right) is needed because of the fact that
we are testing if the population mean is more than the null mean.

### **Performing a student t-test:**
```{r}
t.test(w,mu=100,alternative="greater")
```
The t-test shows us that the p-value is 0.2977, which is different to
the 0.2910448 that I got from calculating the p-value myself. This
is because of the fact that the sample is too small which causes
the uncertainty to be higher.

### **95% and 99% confidence intervals**
```{r}
t.test(w,mu=100,conf.level=0.95,alt="g")
```

```{r}
t.test(w,mu=100,conf.level=0.99,alt="g")
```
The 95 percent confidence interval is 99.13757 and the
99 percent confidence interval is 98.4731.

# **Hypothesis Testing: Binomial Distribution**
### **Probability that one flower is fertilized:**
```{r}
dbinom(1, 7, 0.4)
```

### **Probability that three or more flowers are fertilized:**
```{r}
1 - pbinom(2, 7, 0.4)
```

### **A sensible null hypothesis:**
A null hypothesis would be that H~0~ = 5 since the scientists
found a plant near a beehive to have 5 fertilized flowers.

### **P-value and "more extreme" in this context**
The p-value is defined as the likelihood of 
receiving the observation or a more extreme observation 
if the null hypothesis is true.
More extreme in this context means higher or lower than the
sample mean.

### **One-sided or Two-sided:**
I think a two-sided test is needed because it's not clear
whether we can expect to see an increase or decrease in
the actual observed mean.

### **Calculating the p-value and its significance:**
```{r}
mu = 7*0.4
sigma = sqrt(7*0.4*(1 - 0.4))
SEM = sigma/sqrt(7)
1 - pnorm(mu, 5, SEM)
```
Since the p-value that I got was 0.9994665, I don't think the p-value
is significant because it's not low enough to reject the null hypothesis.
