---
title: "20121865_FishersTest"
author: "Louise Branzuela"
date: '2022-05-28'
output: pdf_document
---
# **Question 1**
### **Null Hypothesis**
A sensible null hypothesis would be that the OS that a student uses
is independent of their course choices. Therefore the null hypothesis is 
H0: (Mac|Art) = (Mac|Science). P-value is a probability that's
calculated under the assumption that the null hypothesis is true.

### **"More Extreme"**
"More extreme" in this case would mean larger than the observed sample mean. 

### **Fisher's test**
```{r}
M <- matrix(c(10,2,11,7),2,2)
dimnames(M) <- list(OS=c("windows","mac"),major=c("science","arts"))
fisher.test(M, alternative="greater")
```
Since the p-value is 0.1869, we fail to reject the null hypothesis
because of the fact that the p-value is greater than 0.05.

\newpage

# **Question 2**
### **Null Hypothesis**
A sensible null hypothesis would be that Vitamin C does not help
in curing the common cold.

### **P-value and "more extreme"**
The p-value is a probability that's calculated under the assumption that
the null hypothesis is true. In this case, "more extreme" means larger
than the sample mean. If the p-value is significantly less than the sample mean,
then that means it is in favour of the alternative hypothesis.

### **One-sided or two-sided**
A one-sided test(right) is needed because of the fact that
we are testing if the population mean is more than the null mean.

### **Estimation of Probabilities**
The probability of a person who was given vitamin C and being cured:
P(Cured|Vitamin C) = P(Cured and Vitamin C) / P(Vitamin C)
```{r}
5/7
```
The probability of a person who was given the placebo and being cured:
P(Cured|Placebo) = P(Cured and Placebo) / P(Placebo)
```{r}
4/13
```

\newpage

### **Fisher's test**
```{r}
N <- matrix(c(5,4,2,9),2,2)
dimnames(N) <- list(
  given=c("vitamin C","placebo"),state=c("cured","cold")
)
fisher.test(N, alternative="greater")
```
Since the p-value of the dataset is 0.1018, we fail to reject
the null hypothesis because of the fact that the p-value is
greater than 0.05.

# **Question 3**
### **P(3 White Balls and 1 Black Ball)**
```{r}
dhyper(3, 3, 5, 4)
```

### **P(2 White Balls and 2 Black Ball)**
```{r}
dhyper(2, 3, 5, 4)
```

### **P(1 White Balls and 3 Black Ball)**
```{r}
dhyper(1, 3, 5, 4)
```

\newpage

### **P(0 White Balls and 4 Black Ball)**
```{r}
dhyper(0, 3, 5, 4)
```

### **Checking if the probabilities sum to 1**
```{r}
sum(dhyper(0:3, 3, 5, 4))
```


# **Question 4**
Fisher's exact test refers to the evaluation of non-random
relationships between two sets of data. In this data,
Fisher's test was used to carry out the frequency data for the results
of the three AUT students and four UoA students to estimate non-randomness.

```{r}
fisher.test(diag(3:4))
```

Since the p-value from the data is 0.02857, we reject the null
hypothesis which means that the test results from the university students
were not random.

\newpage

# **Question 5**
### **Example**
For my example, I'm going to be determining if there's a statistically
significant association between being a computer science major and being a 
gamer. Being a computer science major can only be a yes or a no, 
and being a gamer can only be a yes or no. The two variables presented are 
both qualitative.

### **Contingency Table**
```{r}
O <- matrix(c(9,3,1,6),2,2)
dimnames(O) <- list(
  Major=c("CS Major","Not CS Major"),Gamer=c("Gamer","Not Gamer")
)
O
```

### **Null Hypothesis**
A sensible null hypothesis would be that being a gamer or not is independent
of whether you are a computer science major or not. 
Therefore the null hypothesis is H0: (Gamer|Not CS Major) = (Gamer|CS Major).
The alternative or "more extreme" in this case would mean that the p-value is
larger than the observed sample mean. 

### **One-sided or Two-sided**
A one-sided test(right) is needed because of the fact that
we are testing if the population mean is more than the null mean.

\newpage

### **Fisher's test**
```{r}
fisher.test(O, alternative="greater")
```
Since the p-value from the data is 0.01739, we reject the null
hypothesis which means that being a CS Major and being a gamer at the
same time is not random and is highly likely to be the case that you are
a gamer if you study Computer Science.
